---
title: 'Your Title Here'
author: "Don Francisco"
date: "November 1, 2020"
output: pdf_document
csl: bioinformatics.csl
bibliography: references.bib
params:
  gff_file_path: "/data/sars_vcf_analysis/02_genome_reference/sars_refgenome_annotation.gff"
  vcf_dir_path: "data/11_vcf_output_for_R"
  sra_runtable_path: "data/00_sra_runtable/SraRunTable_PRJNA659293.txt"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      error = TRUE)
```

# Background and Overview

This is a report on SARS-CoV-2, including some variant analysis [@koyama2020variant].

# Methods

## Subsections are ok too

# Results and Discussion

```{r load-packages-and-functions}
library("vcfR")
library("ggplot2")
library("dplyr")
library("ggthemes")

# this loads all function files in the code/functions directory
function_paths <- list.files("code/functions", full.names = TRUE)
purrr::walk(function_paths, source)
```

```{r load-vcf-data}
# load in, tidy, and stack all of the VCF files in the target directory
# these are the output of the bash pipeline
stacked_vcfs <- parse_tidy_and_stack_vcfs(
  vcf_dir_path = params$vcf_dir_path)

# load in the gff file with genome annotations for SARS-CoV-2
gff <- read_gff(gff_file_path = params$gff_file_path)

# pull out just the named genes from the annotation file
gene_table <- extract_genes_from_gff(annotation_object = gff)

# combine the stacked vcf table with the gene names and the metadata from
# the sra runtable file
vcf_with_metadata <- add_genes_metadata_to_vcfstack(
  sra_runtable_path = params$sra_runtable_path,
  stacked_vcf = stacked_vcfs,
  cleaned_genes_table = gene_table)
```

# Figures

```{r sample-plot}
# create a plot of unique SNP locations within each gene across all samples
vcf_with_metadata %>%
  filter(!is.na(gene)) %>% # get rid of SNPs not in gene regions
  group_by(gene, pos) %>%
  tally() %>% # this gives a column n for the number of gene by position
  group_by(gene) %>%
  tally() %>% # this collapses that down to the number of unique SNP locations
  ggplot(aes(x = gene,
             y = n)) +
    geom_col() +
    labs(title = "Count of distinct SNPs in Named SARS-CoV-2 Genes",
         x = "Gene Name") +
  theme_few() # get rid of the grey background
```

**Figure 1**: N and S genes have more unique SNPs in the set of samples analyzed.

# Tables

```{r example-table}
# An example table to show the length of each gene using its start and end
gene_table %>%
  mutate(length = end - start) %>%
  select(gene_name, start, end, length) %>%
  knitr::kable(col.names = c("Gene Name",
                             "Start",
                             "End",
                             "Length"))
```

**Table 1**: Gene names, locations, and lengths in the SARS-CoV-2 genome. Higher SNP counts in the S and N genes may be related to the larger size of these genes.

# Sources Cited
